Search-Based Test Data Generation for JavaScript Functions that Interact with the DOM \thanks{Identify applicable funding agency here. If none, delete this.}

### 2018-06-29 doc/issre2018/graph-generation-time.tex

\documentclass[sigconf,review]{acmart}

\usepackage[inline]{enumitem}
\usepackage{tablefootnote}
\usepackage{syntax}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{booktabs} % For formal tables
\usepackage{tikz}
\usetikzlibrary{arrows,shadows,positioning,decorations.pathmorphing}
\usepackage{pgf-umlsd}
\usepackage{tikz-uml}
\usepackage{hyperref}
\usepackage{upquote}
\usepackage{listings}
\usepackage{lsthaskell}
\usepackage{xspace}
\usepackage{enumitem}
\usepackage{mdframed}
\usepackage{algorithmic}
\usepackage{longtable}
\usepackage{subcaption}
\usepackage{pgfplots}
\pgfplotsset{compat=newest}
\usepackage[titlenotnumbered, linesnumbered, ruled]{algorithm2e} % language for algorithm description
\hypersetup{
  colorlinks   = true, %Colours links instead of ugly boxes
  urlcolor     = blue, %Colour for external hyperlinks
  linkcolor    = red, %Colour of internal links
  citecolor    = green %Colour of citations
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\definecolor{ltblue}{rgb}{0,0.4,0.4}
\definecolor{dkblue}{rgb}{0,0.1,0.6}
\definecolor{dkgreen}{rgb}{0,0.4,0}
\definecolor{dkviolet}{rgb}{0.3,0,0.5}
\definecolor{dkred}{rgb}{0.5,0,0}

\definecolor{lightgray}{rgb}{0.95, 0.95, 0.95}
\definecolor{darkgray}{rgb}{0.4, 0.4, 0.4}
%\definecolor{purple}{rgb}{0.65, 0.12, 0.82}
\definecolor{editorGray}{rgb}{0.95, 0.95, 0.95}
\definecolor{editorOcher}{rgb}{1, 0.5, 0} % #FF7F00 -> rgb(239, 169, 0)
\definecolor{editorGreen}{rgb}{0, 0.5, 0} % #007C00 -> rgb(0, 124, 0)
\definecolor{orange}{rgb}{1,0.45,0.13}		
\definecolor{olive}{rgb}{0.17,0.59,0.20}
\definecolor{brown}{rgb}{0.69,0.31,0.31}
\definecolor{purple}{rgb}{0.38,0.18,0.81}
\definecolor{lightblue}{rgb}{0.1,0.57,0.7}
\definecolor{lightred}{rgb}{1,0.4,0.5}
% CSS
\lstdefinelanguage{CSS}{
  keywords={color,background-image:,margin,padding,font,weight,display,position,top,left,right,bottom,list,style,border,size,white,space,min,width, transition:, transform:, transition-property, transition-duration, transition-timing-function},	
  sensitive=true,
  morecomment=[l]{//},
  morecomment=[s]{/*}{*/},
  morestring=[b]',
  morestring=[b]",
  alsoletter={:},
  alsodigit={-}
}

% JavaScript
\lstdefinelanguage{JavaScript}{
  morekeywords={typeof, new, true, false, catch, function, return, null, catch, switch, var, if, in, while, do, else, case, break, for},
  morecomment=[s]{/*}{*/},
  morecomment=[l]//,
  morestring=[b]",
  morestring=[b]'
}

\lstdefinelanguage{HTML5}{
  language=html,
  sensitive=true,	
  alsoletter={<>=-},	
  morecomment=[s]{<!-}{-->},
  morestring=[b]'
  tag=[s],
  otherkeywords={
  % General
  >,
  % Standard tags
	<!DOCTYPE,
  </html, <html, <head, <title, </title, <style, </style, <link, </head, <meta, />,
	% body
	</body, <body,
	% Divs
	</div, <div, </div>, 
	% Paragraphs
	</p, <p, </p>,
	% scripts
	</script, <script,
  % More tags...
  <canvas, /canvas>, <svg, <rect, <animateTransform, </rect>, </svg>, <video, <source, <iframe, </iframe>, </video>, <image, </image>, <header, </header, <article, </article
  },
  ndkeywords={
  % General
  =,
  % HTML attributes
  charset=, src=, id=, width=, height=, style=, type=, rel=, href=,
  % SVG attributes
  fill=, attributeName=, begin=, dur=, from=, to=, poster=, controls=, x=, y=, repeatCount=, xlink:href=,
  % properties
  margin:, padding:, background-image:, border:, top:, left:, position:, width:, height:, margin-top:, margin-bottom:, font-size:, line-height:,
	% CSS3 properties
  transform:, -moz-transform:, -webkit-transform:,
  animation:, -webkit-animation:,
  transition:,  transition-duration:, transition-property:, transition-timing-function:,
  }
}

\lstdefinelanguage{HTML5}{
  language=html,
  tagstyle=\color{blue},
  markfirstintag=true,
  sensitive=true,
  alsoletter={<>=-},	
  morecomment=[s]{<!--}{-->},
  % tag=[s],
  %% morestring=[b]'',
  morestring=[b]',
  ndkeywords={
    % General
    =,
    % HTML attributes
    id=, class= 
  }
}

\lstdefinestyle{htmlcssjs} {%
  % General design
  backgroundcolor=\color{editorGray},
  basicstyle={\scriptsize\ttfamily},   
  frame=tb,
  % line-numbers
  xleftmargin={0.75cm},
  numbers=left,
  stepnumber=1,
  firstnumber=1,
  numberfirstline=true,	
  % Code design
  identifierstyle=\color{black},
  keywordstyle=\color{blue}\bfseries,
  ndkeywordstyle=\color{editorGreen}\bfseries,
  stringstyle=\color{editorOcher}\ttfamily,
  commentstyle=\color{brown}\ttfamily,
  % Code
  language=JavaScript,
  alsolanguage=HTML5,
  alsodigit={.:;},	
  tabsize=2,
  showtabs=false,
  showspaces=false,
  showstringspaces=false,
  extendedchars=true,
  breaklines=true,
  escapechar=|c
}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\listfiles % used to output the versions of all files being used by the late

\newcommand{\Server}{\emph{Server}\xspace}
\newcommand{\Client}{\emph{Client}\xspace}
\newcommand{\User}{\emph{User}\xspace}
\newcommand{\Validator}{\emph{Validator}\xspace}
\newcommand{\thenBr}{\text{then}}
\newcommand{\elseBr}{\text{else}}
\newcommand{\inFor}{\text{inFor}}
\newcommand{\inWhile}{\text{while}}
\newcommand{\un}[1]{\underline{#1}}
\newcommand{\Random}{\mathbb{R}}
\newcommand{\Genetic}{\mathbb{G}}
\newcommand{\RGenetic}{\mathbb{G}_{10}}
\newcommand{\Confix}{\emph{Confix}}
\newcommand{\Jalangi}{\emph{Jalangi}}
\newcommand{\Jedi}{\emph{JEDI}}

% Copyright
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
\setcopyright{rightsretained}
%\setcopyright{usgov}
%\setcopyright{usgovmixed}
%\setcopyright{cagov}
%\setcopyright{cagovmixed}

% DOI
%\acmDOI{10.475/123_4}

% ISBN
%\acmISBN{123-4567-24-567/08/06}

%Conference
\acmConference[ISSTA 2018]{ACM SIGSOFT International Symposium on Software Testing and Analysis}{16-22 July, 2018}{Amsterdam, The Netherlands}
%% \acmYear{1997}                  
%% \copyrightyear{2016}

%% \acmPrice{15.00}

\begin{document}
\title[Search-Based Test Data Generation for JavaScript Functions...]{Search-Based Test Data Generation for JavaScript Functions that Interact with the DOM}

\author{Alexander Elyasov}
\affiliation{%
  \institution{Utrecht University, The Netherlands} 
}
\email{a.elyasov@uu.nl}

\author{Wishnu Prasetya}
\affiliation{%
  \institution{Utrecht University, The Netherlands}
}
\email{s.w.b.prasetya@uu.nl}

\author{Jurriaan Hage}
\affiliation{%
  \institution{Utrecht University, The Netherlands}
}
\email{j.hage@uu.nl}

\begin{abstract}
The popularity of JavaScript is continuously growing. Together with HTML and CSS, it forms the core technology stack for web development. Because of the dynamic nature and complex interplay with HTML, JS applications are often error-prone and vulnerable. Despite the active research efforts to devise intricate static and dynamic analyses to facilitate JS testing, the problem of test data generation for JS applications interacting with the DOM has not yet been addressed. In this paper, we present a novel \textbf{J}avascript \textbf{E}volution-based testing framework with \textbf{D}OM as an \textbf{I}nput, called \emph{JEDI}. In order to reach a target branch, it applies genetic search for relevant input parameters of the JS function including the global DOM state. We conducted an empirical evaluation to study the effectiveness and efficiency of our testing framework. It shows that the \emph{genetic with restart} algorithm, proposed in this paper, is able to achieve complete branch coverage for all experimental subjects, taking on average 19 seconds per branch.
\end{abstract}

%
% The code below should be generated by the tool at
% http://dl.acm.org/ccs.cfm
% Please copy and paste the code instead of the example below. 
%
\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011074.10011099.10011102.10011103</concept_id>
<concept_desc>Software and its engineering~Software testing and debugging</concept_desc>
<concept_significance>300</concept_significance>
</concept>
<concept>
<concept_id>10011007.10011074.10011784</concept_id>
<concept_desc>Software and its engineering~Search-based software engineering</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[300]{Software and its engineering~Software testing and debugging}
\ccsdesc[300]{Software and its engineering~Search-based software engineering}

\keywords{JavaScript, search-based testing, branch coverage}

\maketitle

%% ------------------------------------------------------------------------
\section{Introduction}
\label{sec.intro}
%% ------------------------------------------------------------------------

According to the \emph{StackOverflow} Developer Survey 2018 \cite{stackoverflow2018}, JavaScript has been recognized for the sixth time in row as the most commonly used programming language. It is also the most popular language on \emph{GitHub}~\cite{guthub2017} based on the number of opened pull requests (2.3 million). Such an ever increasing JS success can be explained by its native browsers support, and the rise in popularity of the JS run-time environment \emph{Node.js}. On the other hand, due to of implicit type conversion, lack of static types, and extensive support of dynamic language constructs~\cite{richards2010analysis}, JS applications are often error-prone~\cite{frolin:TSE16}, hard to analyze~\cite{andreasen2017survey, sun2017analysis} and test~\cite{mesbah2015advances}. Thus, it is not surprising that once in a while critical bugs manage to break through~\cite{bugstories2017} into the web applications by Facebook, Twitter etc.

Systematic unit testing is basic and, at the same time, the most effective technique for improving software quality. In practice, there are numerous of JS frameworks --- Mocha, Jasmine, Jest, Enzyme, Tape, Ava, Karma, QUnit etc. --- intended to assist developers with the automation of various unit testing activities such as design, execution and monitoring. Despite the apparent abundance of the JS testing tools, the developers' satisfaction rate is still relative low~\cite{stateJS2017} (3.2 out of 5), suggesting the need for improvements. One of such areas is \emph{test data generation}. It pursues the goal of maximizing the test coverage by exercising the function under test (FUT) with different input parameters. Moreover, if a FUT interacts with the DOM of a web application, a \emph{test fixture} is expected as an additional input.

Generally, the problem of test data generation has been under an active study for the past three decades. In particular for JS, researchers have proposed several solutions based on: concolic execution (\emph{Jalangi}~\cite{sen2013jalangi}, \emph{Confix}~\cite{amin:ase15}), random generation (\emph{JSContest}~\cite{heidegger2010contract}), and the analysis of the whole web application (\emph{Artemis}~\cite{artemis2011}, Crawljax~\cite{mesbah2012crawling}). Out of those tools, only \emph{Confix} tries to construct a sufficient DOM fixture by translating native browser API calls into logical constraints. However, in the empirical evaluation~\cite{amin:ase15}, Confix only reached about 50\% in \emph{branch coverage} on the selected subjects. At the same time, search-based techniques have shown prominent results in achieving high branch coverage for imperative~\cite{wegener2001evolutionary}, object-oriented~\cite{fraser2011evosuite} and dynamic languages~\cite{irawan2016test, wibowo2015unit}, as well as for web~\cite{alshahwan2011automated} and mobile~\cite{mao2016sapienz} applications. But to the best of our knowledge, the problem of test data generation has never been tackled for JS with the help of the search-based approach. In this paper we are breaching this gap by introducing a \textbf{J}avascript \textbf{E}volution-based testing framework with \textbf{D}OM as an \textbf{I}nput, called \emph{JEDI}. We can summarize our contributions as follows: 
\begin{enumerate}[leftmargin=5mm]
\item The \emph{JEDI} framework is a novel JS unit testing tool using search-based techniques for generating test data. Notably, our testing framework is able to generate an arbitrary DOM input (both tags and attributes), which is syntactically valid and compliant with the recent HTML5 specification.   
\item Our test generation algorithm, called ``genetic with restart'', is trying to escape search plateau by restarting GA with a new target. Because the full-fledged JS data flow analysis~\cite{jang2009points} is hard to reach, the new target is chosen purely based on the preceding control flow behaviour in contrast with more popular techniques~\cite{ferguson1996chaining}.  
\item Based on the case studies found in two related works~\emph{Confix}~\cite{amin:ase15} and \emph{TAJS}~\cite{dom2011}, we conducted an empirical validation of our framework and performed a significance study of the results. The validation has shown the effectiveness of the framework in covering target branches with a reasonable efficiency. Moreover, we showed that on the selected subjects \emph{JEDI} outperforms \emph{Confix} ....  
\end{enumerate}

The paper is structured as follows: In Section~\ref{sec.example}, we present a motivating example illustrating the challenges of generating test data for JS. The architecture of the \emph{JEDI} testing framework is explained in Section~\ref{sec.framework}. Section~\ref{sec.evaluation} presents empirical validation and indicates the validity threats. Related and future work are discussed in Section~\ref{sec:related.work}. Section~\ref{sec:concl} concludes the paper.

%% The state of JS 2017 survey~\footnote{\url{https://stateofjs.com/}} contains an extensive list of testing frameworks including Mocha, Jasmine, Jest, Enzyme, Tape, Ava, Karma and QUnit. These frameworks support testing at various levels: unit, integration and functional. The survey also indicated a relatively low satisfaction rate of the current state of JS testing tools, scoring only 3.2 out of 5. As result, new testing frameworks such as Cypress and Storybook are constantly appearing to improve the situation. But despite this effort, test data generation continues to be poorly addressed by any of the widely used testing frameworks. The research community has actively been working towards making progress in this direction introducing such tools as Artemis~\cite{artemis2011}, JSConTest~\cite{heidegger2010contract}, Crawljax~\cite{mesbah2012crawling}, JSeft~\cite{mirshokraie2015jseft} and Jalangi~\cite{sen2013jalangi}. A recent study of client-side JS bugs~\cite{ocariza2017study} has shown that the majority (68\%) of web application faults are DOM-related and introduced in the JS code, and not in other components such as HTML or CSS. This fact indicates that yet a better testing approach is needed when it comes to unit testing of JS.

%% --------------------------------------------------------------------
\section{Motivating Example}
\label{sec.example}
%% --------------------------------------------------------------------

\begin{figure}[t]
  \begin{lstlisting}[style=htmlcssjs,language=JavaScript]
/*t dom */
function isGameFinished() {
  var obj = document.getElementById('sudoku'); |c \label{isGameFinished.getSudoku} |c
  var subDivs = obj.getElementsByTagName('DIV'); |c \label{isGameFinished.getDivs} |c
  var allOk = true;
  for (var no = 0; no < subDivs.length; no++) { |c \label{isGameFinished.inFor.begin} |c
    if (subDivs[no].className.indexOf('square') >= 0 |c \label{isGameFinished.if1.begin} |c 
        && !subDivs[no].style.backgroundColor) { 
      var spans = subDivs[no].getElementsByTagName('SPAN');
      if (spans[0].innerHTML != spans[1].innerHTML) { |c \label{isGameFinished.if2.begin} |c
        allOk = false; //target |c \label{isGameFinished.unfinished} |c
        break;}}}} |c \label{isGameFinished.outFor} |c
\end{lstlisting}
  \caption{JS function \texttt{isGameFinished.js} from sudoku}
  \label{code.isGameFinished}
\end{figure}

\begin{figure}[t]
  \begin{lstlisting}[style=htmlcssjs, language=HTML5]
<!-- T1: (5,5) -->            <!-- T2: (5,6), (7,15) -->
<html>                        <html>
 <body>                        <body>
  <div id='sudoku'>             <div id='sudoku'>
                                 <div></div>                           
  </div>                        </div> 
 </body>                       </body>    
</html>                       </html>

<!-- T3: (7,8), (10,14) -->   <!-- T4: (7,8), (10,11) -->
<html>                        <html>
 <body>                        <body>
  <div id='sudoku'>             <div id='sudoku'>
   <div class='square'>          <div class='square'>
    <span></span>                 <span></span> 
    <span></span>                 <span>TEST</span>
   </div>                        </div> 
  </div>                        </div>
 </body>                       </body>
</html>                       </html>
  \end{lstlisting}
  \caption{Input DOM arguments for \texttt{isGameFinished.js}}
  \label{fig.isGameFinished.tests}
\end{figure}

Let us consider the JS function \texttt{isGameFinished} in Figure~\ref{code.isGameFinished} taken from the web game \emph{sudoku}~\cite{sudoku}. This function validates a sudoku solution given by a user. First, it finds a DOM element representing the game field (line~\ref{isGameFinished.getSudoku}). Then, it collects all the child \emph{divs} (line~\ref{isGameFinished.getDivs}). Each \emph{div} element corresponds to a sudoku input square (line~\ref{isGameFinished.if1.begin}), and consists of two \emph{spans}. The first \emph{span} contains a user input, whereas the second one (hidden by default) stores an expected value for that sudoku square. Inside of a for-loop in lines~\ref{isGameFinished.inFor.begin}-\ref{isGameFinished.outFor}, we iterate over the whole collection of input squares. Once, we find a square with two unequal values (line~\ref{isGameFinished.if2.begin}), the loop is terminated and the game is considered unfinished (line~\ref{isGameFinished.unfinished}) and finished otherwise.

Suppose now that we would like to unit test this function in isolation to the rest of the application. We aim to maximize the branch coverage as a common testing criteria~\cite{zhu1997software}. That is, for each individual branch of the function under test (FUT) \texttt{isGameFinished}, we need to construct a test input that covers the branch and leads to the normal termination of the FUT. A JS function can take \emph{explicit} input arguments (in our case none), but it can also accept \emph{implicit} arguments such as a DOM state. In general, in order to produce a complete test for our function we have to construct both an appropriate DOM and input arguments.

%Our program has the following branches: $(6,15)$, $(6,7)$, $(7,9)$, $(7,14)$, $(10,11)$, $(10,13)$, $(16,17)$, $(16,18)$. Four tests in

Figure~\ref{fig.isGameFinished.tests} presents four tests that together provide the full coverage of the FUT. Let us take a look at how to construct the \texttt{T4} test, which covers the branch $(10,11)$, by following the structure of the FUT. Line~\ref{isGameFinished.getSudoku} expects the input DOM to have an element with an id \textquotesingle\texttt{sudoku}\textquotesingle. In order to enter the for-loop in line~\ref{isGameFinished.inFor.begin}, that element should have at least one child \emph{div} tag. The first if-condition in line~\ref{isGameFinished.if1.begin} requires this \emph{div} to be of the class \textquotesingle\texttt{square}\textquotesingle\ and also have no background color. The second if-condition (line~\ref{isGameFinished.if2.begin}) expects that the \emph{div} consists of two \emph{span} elements whose \emph{innerHtml} values are not equal. The resulting test \texttt{T4} in Figure~\ref{fig.isGameFinished.tests} meets all the above conditions.

As we have just observed, even for such a relatively simple example, the test data generation can be far from trivial procedure because it requires the deep understanding of the program semantics. In practice, web developers often have to deal with a large number of the JS frameworks and libraries freely available on the GitHub whose code is commonly untested and poorly documented. At the same time, unit testing is the most fundamental level of testing and the easiest to implement. It saves the developers from unexpected regressions and often documents the desired behavior of a program.

%% --------------------------------------------------------------------
\section{Test Generation Framework}
\label{sec.framework}
%% --------------------------------------------------------------------

\begin{figure}[t!]
  \centering 
\tikzset{
  % add this style to all tikzpicture environments
  every picture/.append style={
    % enable scaling of nodes
    transform shape,
    % set scale factor
    scale=1
  }
}
  \begin{sequencediagram}[font=\scriptsize]

    \renewcommand\unitfactor{0.35}
    \newinst{user}{\textbf{User}}
    \newinst{server}{\textbf{Server}}
    \newinst[1]{validator}{\textbf{Validator}}
    \newinst{client}{\textbf{Client}}

    \begin{call}{user}{fun.js + config}{server}{\textbf{end}}
      \begin{call}{server}{Initialize(fun.js)}{server}{(cfg, branches, cinfo, ifun, sig)}
      \end{call}

      \begin{call}{server}{POST /init \{ifun, sig\}}{client}{status}
      \end{call}

      \begin{call}{server}{Select(branches)}{user}{$\text{branches}'$}       
      \end{call}

      \begin{sdblock}{\textbf{Genetic Loop}}{\scriptsize \hspace{8mm}$\text{branch} \in \text{branches}'$}
        \begin{sdblock}{\textbf{Random Loop}}{\scriptsize errors > 0}
          \begin{call}{server}{pop = Random(cinfo, config)}{validator}{errors}
          \end{call}
          \prelevel
        \end{sdblock}

        \begin{sdblock}{\textbf{Fitness Loop}}{\scriptsize (fitness $\neq$ 0) $\vee$ gen.limit}

          \begin{sdblock}{\textbf{Evaluation Loop}}{\scriptsize p $\in$ pop}
            \begin{call}{server}{POST /genetic \{p\}}{client}{response = ifun(p: sig)}
            \end{call}

            \begin{callself}{server}{Evaluate(branch, cfg, response)}{fitness}
            \end{callself}
            \prelevel
          \end{sdblock}

          \begin{sdblock}{\textbf{Crossover Loop}}{\scriptsize errors > 0}
            \begin{call}{server}{cross = Crossover(cinfo, pop)}{validator}{errors}
            \end{call}
            \prelevel
          \end{sdblock}

          \begin{sdblock}{\textbf{Mutation Loop}}{\scriptsize errors > 0}
            \begin{call}{server}{mut = Mutation(cinfo, pop)}{validator}{errors}
            \end{call}
            \prelevel
          \end{sdblock}

          \begin{callself}{server}{pop = cross $\cup$ mut}{pop}
          \end{callself}
          \prelevel
        \end{sdblock}
        \mess{server}{best fitness}{user}

      \end{sdblock}
    \end{call}
  \end{sequencediagram}
  \caption{The architecture of the test framework}
  \label{fig.framework.architect} 
\end{figure}

Given a JS function $f$ and a set of branches $\mathbb{B}_f$, the ultimate goal of the testing framework is to find input arguments for $f$ that cover all branches from $\mathbb{B}_f$. Our JS testing framework, called \emph{JEDI}, is a client-server application, which uses under the hood a genetic engine, for test generation. The general workflow of the framework is depicted as a sequence diagram in Figure~\ref{fig.framework.architect}. The diagram consists of the four interacting components \User, \Server, \Validator, and \Client. The \Server is the key component of the  framework, responsible for the \emph{Initialization} and \emph{Genetic Phases}, covered in detail in Section~\ref{sub.sec.init.phase} and \ref{sub.sec.genetic.phase} respectively.

Each newly constructed HTML document has to be assessed by the \Validator~\cite{htmlvalidator} that checks its syntactic correctness and reports back any identified errors. The document generation process is repeated by the \Server until there are no errors revealed. \Client is a \emph{Node.js}~\cite{nodejs} application that is mainly responsible for the execution of the JS code. We use the \emph{jsdom}~\cite{jsdom} library to model the virtual DOM and simulate the native browser API calls.

\begin{algorithm}[t!]
  \caption{Initialization Phase}
  \label{alg.init}
  \algsetup{linenosize=\tiny}
  \scriptsize
  \DontPrintSemicolon
  \SetAlgoVlined
  \SetKwInOut{Input}{Input}
  \SetKwInOut{Output}{Output}
  \SetKwInOut{Require}{Require}
  \SetKwProg{Fn}{Function}{}{}
  \Input{JS file $fun.js$ with FUT and type annotation }
  \Output{Tuple $(cfg, branches, cinfo, ifun, sig)$}
  \Fn{$Initialize(fun.js)$}{

    $(ast, sig) \longleftarrow ParseFuncAndSig(fun.js)$\label{alg.init.parse.fun.sig}\;
    $cinfo \longleftarrow GetConstantInfo(ast)$\label{alg.init.collect.info}\;
    $nast \longleftarrow NormalizeAST(ast)$\label{alg.init.transform.ast}\;
    $ifun \longleftarrow Instrument(nast)$\label{alg.init.instr}\;
    $cfg \longleftarrow BuildCFG(nast)$\label{alg.init.build.cfg}\;
    $branches \longleftarrow GetBranches(cfg)$\label{alg.init.get.branches}\;
    \Return{$(cfg, branches, cinfo, ifun, sig)$}
    }
\end{algorithm}

Let us take a closer look at the sequence of interactions in Figure~\ref{fig.framework.architect}. As an input for our testing framework, \User has to provide a JS file \emph{fun.js} that contains a FUT \emph{fun} annotated with the type signature. Additionally, \User specifies a configuration file \emph{config} that sets various control parameters for the random and genetic phases, logging etc. Experimentally, we have found an optimal configuration that was consistently used in the experimental evaluation in Section~\ref{sec.evaluation}. Once all the data are processed by the \Server in \emph{Initialization Phase}, we get back the following information about \emph{fun}: the control flow graph (\emph{cfg}), the complete list of its branches (\emph{branches}), the constant info (\emph{cinfo}), the instrumented version \emph{fun} (\emph{ifun}), and the signature (\emph{sig}). Then, the \Server sends a \texttt{POST} request with \emph{ifun} and \emph{sig} to the \Client. Depending on the configured execution mode, the \Server either asks the \User to select a \emph{target} branch or covers all branches one by one in a fixed order. For each branch, our framework triggers the \emph{Genetic Loop (Phase)}. Inside of this phase, the framework begins with a randomly generated population (\emph{Random Loop}), and continues to evolve that population (\emph{Fitness Loop}) until either a perfect entity is found or a termination criteria is achieved, e.g. the limit on the number of generation attempts. Any HTML document produced during the genetic phase has to be first approved by the \Validator. Each candidate in the population has to be evaluated in \emph{Evaluation Loop} by \Client. It matches the candidate solution $p$ with the type signature \emph{sig}, and invokes the instrumented JS function as \emph{ifun(p)}. The execution trace along with some auxiliary information is sent back to the \Server, where the final \emph{fitness value} is computed. A new population is constructed by combining the results of the \emph{Crossover} and \emph{Mutation Loops}. At the end of the \emph{Fitness Loop}, the candidate with the ``best'' fitness value is reported back to the \User. Once all the branches are covered, the algorithm terminates. Below, we elaborate on the details of each individual phase of our test generation framework.

\subsection{Initialization Phase}
\label{sub.sec.init.phase}

During the \emph{Initialization Phase} we analyze the FUT in order to facilitate the subsequent genetic testing. The main steps of this phase are described in Algorithm~\ref{alg.init}. Given a FUT \emph{fun.js} as an input, we parse the function and its type signature in line~\ref{alg.init.parse.fun.sig}. In order to maximize the chances of generating the fittest candidate, in line~\ref{alg.init.collect.info} we perform static analysis of the FUT with the purpose to collect information about constants, e.g. numeric and string literals. Additionally, we monitor the calls to such DOM API methods as \texttt{getElementsByTagName}, \texttt{getElementById}, and \texttt{getElementsByClassName}. These methods provide the information about \emph{tags}, \emph{ids}, \emph{classes} and \emph{names} referred from the FUT. The next step is to normalize the FUT before applying any instrumentation (line~\ref{alg.init.transform.ast}), e.g. transform \emph{if-then} branches into \emph{if-then-else}. We instrument the FUT (line~\ref{alg.init.instr}) for run-time data collection about the execution trace, \emph{the approach level}, \emph{the branch and loop distances} (explained later), and the constants. Upon the \emph{ifun} execution this information is gathered and fed back to the fitness evaluation procedure. Finally, we construct the control-flow graph (line~\ref{alg.init.build.cfg}) and identify all branches of the FUT (line~\ref{alg.init.get.branches}).

\subsubsection{Supported Types}
\label{sub.sec.sup.types}

\begin{figure}[t!]
\setlength{\grammarparsep}{3pt}
\footnotesize
\begin{grammar}
<JS Type Annotation> ::= /*t <Signature> */

<Signature> ::= \texttt{dom} : <Type Signature> | <Type Signature>

<Type Signature> ::= <Type> | <Type> : <Type Signature>

<Type> ::= <Void Type> | <Primitive Type> | <Array Type>

<Primitive Type> ::= \texttt{bool} | \texttt{int} | \texttt{float} | \texttt{string}

<Array Type> ::= [ <Primitive Type> ] | [ <Array Type> ]
\end{grammar}
\caption{Grammar of supported JS type annotations}
\label{fig.js.type.annot}
\end{figure}

We assume the FUT is equipped with the type signature. The legitimacy of such a requirement can be confirmed by the popularity of \emph{TypeScript}~\cite{typescript}, and advances in the JS type inference \emph{Flow}~\cite{flow}. The supported grammar of type annotations is shown in Figure~\ref{fig.js.type.annot}. Each type annotation is a signature embedded into a specialized JS type comment (\texttt{/*t...*/}). We distinguish between two kinds of signatures: with the reference to the global DOM and ``ordinary'' JS type signatures. The former kind expects the \texttt{dom} terminal to be in place of the first parameter of the function call. Whereas an ordinary JS signature is a non-empty sequence of the JS types. The type can be \emph{Void}, \emph{Primitive} (\texttt{bool}, \texttt{int}, \texttt{float}, \texttt{string}) or \emph{Array} (recursive homogeneous array type).

\subsubsection{Instrumentation}
\label{sub.sec.instrument}

Figure~\ref{code.isGameFinished.instr} presents the \texttt{isGameFinished} function after instrumentation. The global variable \texttt{trace} in line~\ref{isGameFinished.instr.trace} records the sequence of executed statements of the FUT. We have to trace all the statements because each one, in theory, can rise an exception. The \texttt{loopMap} variable in line~\ref{isGameFinished.instr.loopMap} captures the upper bound for the number for-loop iterations. If the exact bound is unknown, it is set to one. The \texttt{branchDistance} variable, computed for the \emph{problem node}, contains the distance~\cite{tracey1998automated} from the target branch.

%TODO: maybe expand this section with the JS type casting for branch distance computation

\begin{figure}[t]
  \begin{lstlisting}[style=htmlcssjs,language=JavaScript]
function isGameFinished() {
 trace.push(1);var obj = document.getElementById("sudoku"); |c \label{isGameFinished.instr.trace} |c
 trace.push(2);var subDivs = obj.getElementsByTagName("DIV");
 var allOk = true;
 loopMap[3] = function () { |c \label{isGameFinished.instr.loopMap} |c
   var no = 0;
   if (subDivs.length || subDivs.length == 0) {
     return Math.abs(no - subDivs.length);
   }; return 1}();
 trace.push(3);
 for (var no = 0; no < subDivs.length; no++) {
  trace.push(4);
  if (subDivs[no].className.indexOf("square") >= 0 &&
      !subDivs[no].style.backgroundColor) {
   trace.push(5); branchDistance.push({ label: 3, 
    distance: Math.min(
     abs(subDivs[no].className.indexOf("square"),0)+_K_ 
     absZero(subDivs[no].style.backgroundColor))
   });
   trace.push(6);var spans=subDivs[no].getElementsByTagName("SPAN");
   trace.push(7);
   if (spans[0].innerHTML != spans[1].innerHTML) {
    trace.push(8); branchDistance.push({ label: 7, 
     distance: abs(spans[0].innerHTML, spans[1].innerHTML)
    });
    allOk = false; break;
   } else {
   trace.push(9);branchDistance.push({label:7,distance:_K_});
   }
  } else {
   trace.push(10); branchDistance.push({ label: 3,
    distance: 
     abs(subDivs[no].className.indexOf("square"), 0) + 
     absNegZero(subDivs[no].style.backgroundColor)
   });}}
 trace.push(-1);
}
\end{lstlisting}
  \caption{Instrumented version of \texttt{isGameFinished.js} }
  \label{code.isGameFinished.instr}
\end{figure}

\subsection{Genetic Phase}
\label{sub.sec.genetic.phase}

>>>>

  \begin{tikzpicture}[scale=0.9,rotate=-90]

    \begin{axis}[
        every axis plot post/.style={/pgf/number format/fixed},
        ybar,
        ymin=0,ymax=950,
        width=11cm,height=3.3cm,
        bar width=4pt,
        enlarge x limits=0.05,
        legend style={rotate=45,anchor=north west,font=\footnotesize,legend columns=-1,at={(0.05,1.1)}},
        legend cell align={left},
        ylabel={time (sec.)},
        symbolic x coords={helpMe*,isGameFinished*,newGame*,revealAll*,shuffleBoard*,toggleInfo*,isValidCard,isValidVISA,initShields*,brickJiggler*,doPaddlePower*,initBricks*,do\_draw,prob\_gamma},
        xtick=data,
        restrict y to domain*=0:1350, % Cut values off at 14
        visualization depends on=rawy\as\rawy, % Save the unclipped values
        after end axis/.code={ % Draw line indicating break
          \draw [ultra thick, white, decoration={snake, amplitude=1pt}, decorate] (rel axis cs:0,1.05) -- (rel axis cs:1,1.05);
        },
        nodes near coords={%
          \pgfmathprintnumber{\rawy}% Print unclipped values
        },
        every node near coord/.append style={rotate=90,anchor=west,font=\footnotesize},
        % every node near coord/.append style={font=\footnotesize},
        axis lines*=left,
        clip=false,
        x tick label style={rotate=90,anchor=east,font=\footnotesize\bfseries},
        y tick label style={font=\footnotesize}
      ]
      \addplot[fill=blue] coordinates {(helpMe*,270) (isGameFinished*,220) (newGame*,14) (revealAll*,50) (shuffleBoard*,29) (toggleInfo*,6) (isValidCard,514) (isValidVISA,95) (initShields*,5) (brickJiggler*,8) (doPaddlePower*,15)  (initBricks*,2379) (do\_draw,625) (prob\_gamma,710)};
      \addplot[fill=red] coordinates {(helpMe*,47)  (isGameFinished*,30)  (newGame*,9)  (revealAll*,44) (shuffleBoard*,26) (toggleInfo*,7) (isValidCard,8)   (isValidVISA,3)  (initShields*,5) (brickJiggler*,4) (doPaddlePower*,140) (initBricks*,474) (do\_draw,1295) (prob\_gamma,84)};
      \addplot[fill=yellow] coordinates {(helpMe*,56)  (isGameFinished*,43)  (newGame*,10) (revealAll*,53) (shuffleBoard*,29) (toggleInfo*,6) (isValidCard,10)  (isValidVISA,2)  (initShields*,3) (brickJiggler*,3) (doPaddlePower*,16) (initBricks*,415) (do\_draw,334) (prob\_gamma,81)};
      \legend{$\Random$,$\Genetic$,$\RGenetic$}
    \end{axis}
  \end{tikzpicture}
